A flexible framework to prototype, develop and manage ubiquitous applications for intelligent environments

### 2013-10-04  labtempo-ubicomp-projects/UbicompMiddleware/doc/PerCom2014/percom2014_conf.tex
Change the order to the 'paper remain' section

The remainder of this paper is organized as follows. First we present in Section~\ref{sec:overview} an overview of the main concepts used as a basis for the framework's development. In Section~\ref{sec:framework}, we present the framework's overall architecture and main features. We then, in Section~\ref{sec:evaluation}, present a proof of concept demonstrating the feasibility of building UbiApps using the framework and we show examples of applications that explores the key features of the IE. Finally, in Section~\ref{sec:related} we compare our approach with related work and in Section~\ref{sec:conclusion} we conclude this paper with the main remarks.
[refactor this paragraph]

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%              SECTION
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Related Work} \label{sec:related}
% [ubicomp is nice but it is not priceless]
The benefits to users and developers that UbiComp foresees reach beyond the horizon of many imaginative researchers. The possibilities that arise from it surpass the bounds established by keyboards and mice, our standard interfaces. Nonetheless, building such adaptable applications requires much effort from developers that see themselves immersed in a universe of device's specifications and communication technologies, apart from the problem itself that the application must solve. Therefore, many researchers have focused on diminishing those obstacles by providing abstractions, middlewares, services, tools, and other supportive techniques, not only for development but also for prototyping of UbiApps.

@@

%TODO: Gaia
As one of the most famous, the Gaia middleware is designed to facilitate the construction of applications for IE. It consists of a set of core services and a framework for building distributed context aware applications. The Gaia embraces many goals that include the acquisition of context, the maintenance of hardware and software descriptions, mechanisms to find resources and to compose context rules, among others.
%uses operating systems abstractions to define its architecture
%have many goals that include the provision of mechanisms of acquisition of context; composition of rules (using first logic order), 
%%maintenance of hardwares and softwares description and about applications; 
%search mechanism to find resources
%presence and execution status of resources
%interpretation mechanism that gathers and interprets informations, as defined by programmers
%mvc
%The main goals of Gaia's approach are the acquisition of context by applications, the monitoring of entities location, the maintenance of hardwares and softwares descriptions, and 
%Active Spaces
%FIXME: second phrase was copied

@@

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%              SECTION
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Framework Features}
This section approaches the main features provided by the framework, which includes the definition of the RA (Subsection~\ref{subsec:ra}), the management services (Subsection~\ref{subsec:sgar}) and the context interpretation (Subsection~\ref{subsec:ci}). In the next section we will present the current status of the implementation.
% This section approaches the main features provided by the framework, which includes the definition of the RA (Subsection~\ref{subsec:ra}), the management services (Subsection~\ref{subsec:sgar}) and the context interpretation (Subsection~\ref{subsec:ci}). Also, based on the reference implementation, an application was built to provide a tool for the prototyping, management and visualization of applications and resources running on the IE.

@@

    The RA architecture includes two interfaces for input and output. The former is called Context Variable (CV) and is responsible for exposing the RA's context information. The latter is called Operation (OP) and is responsible for exposing features (or actuation services) belonging to RAs, enabling applications to actively interact in the environment.
    % In the previous example, the ``oven is turned on'' and the ``oven temperature in ºC'' are the some of the RA's CVs.

    Consider the simple example of a smart stove where the instantiation of its RA exposes services to get its current context (e.g., oven is turned on, oven temperature in $\,^{\circ}\mathrm{C}$) and services that may change its context (e.g., turn oven off, set temperature as $150\,^{\circ}\mathrm{C}$). Hence, the RA concept meets the first challenge (i), which is the \textit{devices heterogeneity}, since decoupling sensor and actuator from the sensor platforms favors the flexibility. %FIXME: is it right?

@@

\section{SmartAndroid}
% [android and other techs]
%We consolidated the concepts elaborated in the course of this project as the platform \mbox{\textit{SmartAndroid}}\footnote{For more information visit www.tempo.uff.br/smartandroid}. The project includes the implementation of APIs to , as well as prototyping and management interface on the Android platform.

%[David]
The conceptual framework proposed at this work was tested through a reference implementation called \mbox{SmartAndroid} Project\footnote{For more information visit www.tempo.uff.br/smartandroid}. The project includes a programming API to help the developers in constructing UbiApps as well as offers basic management services to provide resource registration, discovery and localization. The SmartAndroid has been developed over Google Android platform as we believe that in a near future it will be widely used in sensors and devices embedded in furnitures and home appliances.
% [we have implemented most of the framework features, thought much more remain to be done]

%proposta de uso de marketplace
% A exposição de dados por meio de \textit{Content Provider} permite a integração de aplicativos externos ao servidor. Estes aplicativos podem ser dispostos em um \textit{marketplace} (http://developer.android.com/distribute/open.html) permitindo que tanto usuários comuns quanto programadores incrementem o ambiente com novas aplicações ubíquas. 
The SmartAndroid middleware contemplates the possility of running multiple applications in the same IE. Thus, focusing on the user and inspired by the Google Play -- which is a application market for Android devices --, we have as a future work the creation of a SmartAndroid Market~\cite{google2013}, where users will be able to easily download and start different applications. Also similar to the Google Play, a Manifest file will require permission for accessing the domains, what will allow users to maintain control over what off-the-shelf applications have permission to know and to do. Within the market, the UbiApp could be tested towards discovering any possible harm that the application could do and any possible inefficiency that could consume too much the energy resources of the host device.
% TODO: talk about market place
% [security concern on store: helps the discovery of malware before the user download]

  \subsection{Location}
  [talk about location]

  \subsection{Communication Support}
  %[talk about stub intantiation calls]
  %TODO: example of CV and OP? Maybe it is worth add some on the text
  To perform any interaction with a RA that is not locally instantiated, the UbiApp must use a stub (also called proxy) that is bound to the referred RA. This stub implements the same interface as the RA, i.e. the same CVs and OPs, which are methods, from a Object-Oriented Programming perspective. With the stub on hands, the application can invoke remote methods, as if they were locally invoked. These remote methods may be to retrieve a CV or to trigger an OP. The Figure~\ref{fig:rpc} represents this direct, or synchronous, communication. The steps to achieve a direct communication, as represented by the figure, are as follows:
  \begin{enumerate}
    \item The UbiApp calls a method from the stub 
    \item The call is forwarded to the RA instance through the network in the second step
    \item This instance can interact with the device (the lamp, in case)
    \item The result or the acknowledge is then returned to the caller
  \end{enumerate}

@@

  \subsection{Context Rules} \label{subsec:rules}
%   [implementation of context rules]
  The SmartAndroid supports the context interpretation by providing mechanisms to create context rules. The context rules life cycle has two major phases: composition and running. The composition phase corresponds to the instantiation of a CI and of an actuator and the binding of both through subscription. The running phase is when the CI and the actuator(s) are already instantiated as the other RAs that feed the IE with context information, and they are able to exchange messages and perform their job.
  %   Support the interpretation of the context framework enables a separation of concerns, where developers abstract implementation of contextual rules and focus on the application logic.

  At the implementation level, a context rule can be described in a JSON structured file (or byte stream) following a specific format. In the first part an expression must be defined setting the logical operat
  To separate the context rules creation from the

%   O IC é encapsulado em um AR, estendendo, assim, suas funcionalidades. Sua criação ocorre quando é definida uma expressão lógica no objeto do IC recém instanciado, segundo um padrão próprio adotado em arquivos tipo JSON, e um \textit{parser} constrói uma estrutura lógica interna, como mostrado na Figura~\ref{fig:inpt_arq}. A estrutura lógica abrigada pelo IC, contém uma estrutura em árvore que representa a expressão e contém não só a lógica entre as VCs das condições, como também os valores atualizados das mesmas. Enquanto em funcionamento, o IC recebe atualizações das VCs, tratando-as em seguida para atualizar sua estrutura lógica. O módulo de avaliação é notificado de mudanças da estrutura lógica e avalia se a expressão se tornou verdadeira ou falsa. Se a expressão alcançou um estado verdadeiro, considerando as possíveis temporizações, notifica os atuadores.

  [explain that both the rule expression and the actuator are written in the JSON]

  [context rules conflicts is an ongoing study that aims at discover a conflict in runtime,]

@@

  %[David]
  The main purpose of the [IPGAP] is provide to developers a hybrid environment where simulated and real resources can interact to each other transparently, enabling their users to construct UbiApps as well as test and evaluate them through a wide variety of scenarios. Thus RAs from sensors, actuators and household appliances can be graphically represented at our tool, along with a schematic map from the intelligent environment so that some of the resources aspects as its context and location can be easily visualized on the map, helping in application debugging.

  Besides, [IPGAP] operate with two different interfaces: the \textit{developer interface}, already mentioned, and the \textit{end-user interface}. In the end-user interface, users in general, i.e. non-technical users, are able to use [IPGAP] to manage their intelligent environment remotely using a mobile device, such as a tablet. Through an intuitive GUI, end-users can not only visualize the state of their home devices but control them. For instance, it's possible turning off a stove burner, change the air-conditioner temperature, and so on. By touching the device icon in the screen, an interface that presents the device's main functions is shown, where the user can operate it as the tool were a remote control.

  % [main figure]
  \begin{figure}[ht]
  \centering
  \includegraphics[width=3.5in]{figures/map-forniture-cv-menu-big-v2.pdf}
  \caption{Management and Prototyping Interface} %[IPGAP]
  \label{fig:ipgap}
  \end{figure}
  %FIXME: convert portuguese text to english

%   [rules composer]
  %TODO Think about a better rule example!!
  In addition, end-users can compose context rules in a hi-level way by graphically combining the CVs from any RA and logical connectives, forming logical expressions as \textit{``if the TV from living room \textbf{is on}, \textbf{and} nobody has been present at the living room for 15 minutes, \textbf{then} turn off that TV''}. To compose this rule the user have to, first of all, select the rule composer mode in the [IPGAP]. Then, he must touch the TV icon at the living room representation on the map, and choose the desired CV -- in that case, the CV that refers if the TV is turned on or not. After that, a dialog box showing all possible logical operators is shown, so that the user can select one of them as the \textbf{AND} operator. The user can repeat this process, adding others CVs and operators until the expression is complete. Finally, the rule can be associated to one or more actions, provided by actuators.
  %FIXME: here ``AND'' is bold, but before it was just between quotation marks

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%              SECTION
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Example Applications} \label{sec:evaluation}
Some applications have been built to explore the potential of the framework for building UbiApps. Although the example applications are simple, and alone does not justify the deploy of the SmartAndroid in the environment, they have allowed us to validate the features provided by the framework and achieved by the SmartAndroid. It also demonstrated the possibility of concomitant running of UbiApps, what similar to an operation system where different processes run separately but in the same system. Withing the SmartAndroid, UbiApps run separately but in the IE.% and can share resources,

% In this section we will elaborate over the SmartLiC application, and next we will briefly talk about the others.
In the subsections that follows we will elaborate over some UbiApps that have been developed, or are still in development, in the course of the project.
% the SmartLiC, the MediaFollowMe, remote control type applications where resources are controlled remotely by another resource (in this case, a smartphone ), among other applications. A future work is to install a system in Ambi home care health, as in ~ \ cite {} carvalho2010sciads.

@@

  The [IPGAP] has allowed us to test this application. We instantiated lamp simulators for each room and an avatar for a user, and also a simulator for day light sensor. Next context rules were created for each room of the house. %Next context rules were created by filling a skeleton of a rule written in json, what caused the instantiation of a CI and actuator for each room.  If the avatar is dragged, in the [IPGAP], from a room and dropped in another, and the specified time passes, the the lamp is turned off.
  \begin{enumerate}
   \item Create a context rule skeleton in JSON
   \item Get the rooms from the house
   \item Get the lamp and presence sensor of each room and fill the rule expression skeleton
   \item Instantiate a CI with the rule expression
   \item Instantiate a actuator to perform the actions declared in the context rule skeleton
   \item Bind the CI and the actuator
   \item Repeat this for each room
  \end{enumerate}

@@

  The SmartLiC allowed us to validate two features of the RLS, which are the discovery of the rooms and the RAs of some specific type are within the room (i.e. the lamps and presence sensors). In this UbiApp we also have created a set of context rules with timers using the context rules parser (including CI and actuators). And the [IPGAP] were well explored in this implementation, since we has shown the triggering of many events in parallel and multiple users cohabiting the same IE.
% 
%   After the start of the context rules, 
%   \begin{enumerate}
%    \item On event received, get the sender, the 
%   \end{enumerate}

@@

  \begin{figure}[ht]
  \centering
  \includegraphics[width=3.5in]{figures/beagle-lamp.pdf}
  \caption{Remote Control} %[IPGAP]
  \label{fig:remotecontrol}
  \end{figure}

@@

  This game was implemented with simulated resources and using the [IPGAP]. The players were represented with avatars and their location were updated with the drag and drop of the avatars over the map.

@@

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%              SECTION
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion} \label{sec:conclusion}
% [framework main features]
In the present paper we have proposed a framework that address some major issues that are found in the construction of ubiquitous applications, both for developers as for end users. We have focused on three challenges, that have guided the development of the framework. For dealing the the heterogeneity of devices present in the intelligent environment we have created the Resource Agent abstraction that encapsulates the mechanisms to interact with physical and also virtual resources, and includes a synchronous and asynchronous communication mechanism. The variety of context information and services is achieved by the definition of Context Interpreters that are responsible for aggregating context information, providing it in a higher level. And for the resource availability issue we have created the [IPGAP], which is a software that helps the prototyping and testing of ubiquitous applications through a easy to use interface, and allows the visualizations of applications running in a environment populated with real and simulated devices.

@@

% [future work]
% [more complex resource discovery]
% [aggregate other security concerns]
% [machine learning and data mining]
% [internet of things - expand to internet]
As future work we plan to extend the framework to allow more complex queries in the resource discovery service, what nowadays is still strict; adapt the communication mechanism for using web services, and get our approach closer to the concept of the internet of things; further research on other security concerns; provide mechanisms to avoid/solve potential conflicts on context rules; aggregate machine learning techniques that will enable a better understanding of user preferences, and will make possible the offering of context rules and other helping services; improve the [IPGAP] and evaluate its usability, using methodologies of human-machine interface; and son on.

>>>>

The remainder of this paper is organized as follows. First we compare our approach with related works in Section~\ref{sec:related}, then in Section~\ref{sec:overview} we present an overview of the main concepts used as a basis for the framework's development. In Section~\ref{sec:framework}, we describe the framework's overall architecture as well as, in Section~\ref{sec:features}, its most important features. Our framework reference implementation as the assumptions, techniques and features related are discussed in Section~\ref{sec:smartandroid}. A proof of concept demonstrating the feasibility of building UbiApps using the framework is presented in Section~\ref{sec:evaluation}, showing examples of applications that explores the key features of the IE. Finally, in Section~\ref{sec:conclusion} we conclude this paper with the main remarks and future works.

%The remainder of this paper is organized as follows. First we present in Section~\ref{sec:overview} an overview of the main concepts used as a basis for the framework's development. In Section~\ref{sec:framework}, we present the framework's overall architecture and main features. We then, in Section~\ref{sec:evaluation}, present a proof of concept demonstrating the feasibility of building UbiApps using the framework and we show examples of applications that explores the key features of the IE. Finally, in Section~\ref{sec:related} we compare our approach with related work and in Section~\ref{sec:conclusion} we conclude this paper with the main remarks.
[refactor this paragraph]

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%              SECTION
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Related Work} \label{sec:related}
% [ubicomp is nice but it is not priceless]
The benefits to users and developers that UbiComp foresees reach beyond the horizon of many imaginative researchers. The possibilities that arise from it surpass the bounds established by common interface devices, such as mice, keyboards and touch screens. Nonetheless, building such adaptable applications requires much effort from developers who see themselves immersed in a universe of device's specifications and communication technologies, apart from the problem itself that the application must solve. Therefore, many researchers have focused on diminishing those obstacles by providing abstractions, middlewares, services, tools, and other supportive techniques, not only for development but also for prototyping of UbiApps.

@@

[best classical references]
\cite{satyanarayanan2001pervasive}
\cite{Park2007simulation}
\cite{ranganathan2003middleware}
\cite{dey2001conceptual}
\cite{helal2005gator}

[others]
\cite{brush2011home}
\cite{edwards2001home}
\cite{garcia2010towards}
\cite{baldauf2007survey}
\cite{gu2005service}
\cite{kindberg2002system}

[only old references]

[get references from this conference]

%TODO: Gaia
Being one pioneer in the field, the Gaia middleware was designed to facilitate the construction of applications for IE.It consists of a set of core services and a framework for building distributed context aware applications. The Gaia embraces many goals that include the acquisition of context, the maintenance of hardware and software descriptions, mechanisms to find resources and to compose context rules, among others.
% FIXME: it really permits to compose context rules or just proposes?

%uses operating systems abstractions to define its architecture
%have many goals that include the provision of mechanisms of acquisition of context; composition of rules (using first logic order), 
%%maintenance of hardwares and softwares description and about applications; 
%search mechanism to find resources
%presence and execution status of resources
%interpretation mechanism that gathers and interprets information, as defined by programmers
%mvc
%The main goals of Gaia's approach are the acquisition of context by applications, the monitoring of entities location, the maintenance of hardwares and softwares descriptions, and 
%Active Spaces
%FIXME: second phrase was copied

@@

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%              SECTION
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% FIXME: Framework Details????????????????????
\section{Framework Features} \label{sec:features}
This section describes details of the main features provided by the framework, which includes the definition of the RA (Subsection~\ref{subsec:ra}), the management services (Subsection~\ref{subsec:sgar}) and the context interpretation (Subsection~\ref{subsec:ci}). In the next section we will present the current status of the implementation.
% This section approaches the main features provided by the framework, which includes the definition of the RA (Subsection~\ref{subsec:ra}), the management services (Subsection~\ref{subsec:sgar}) and the context interpretation (Subsection~\ref{subsec:ci}). Also, based on the reference implementation, an application was built to provide a tool for the prototyping, management and visualization of applications and resources running on the IE.

@@

    The RA architecture includes two options of interface used for different purposes. The former provides access to Context Variables (CV) and is responsible for exposing the RA's context information. The latter is called Operation (OP) and is responsible for exposing internal attributes (reading) or activating internal attributes (writing) belonging to RAs, enabling applications to manage RA's internal state and explicitly interact with the environment.

    % In the previous example, the ``oven is turned on'' and the ``oven temperature in ºC'' are the some of the RA's CVs.

    Consider the simple example of a smart stove where the instantiation of its RA exposes services to get its current context (e.g., oven is turned on, oven temperature in $\,^{\circ}\mathrm{C}$) and services that may change its context (e.g., turn oven off, set temperature as $100\,^{\circ}\mathrm{C}$). Hence, the RA concept meets the first challenge (i), which is the \textit{devices heterogeneity}, since decoupling sensor and actuator from the sensor platforms favors the flexibility. %FIXME: is it right?

@@

\section{SmartAndroid} \label{sec:smartandroid}
% [android and other techs]
%We consolidated the concepts elaborated in the course of this project as the platform \mbox{\textit{SmartAndroid}}\footnote{For more information visit www.tempo.uff.br/smartandroid}. The project includes the implementation of APIs to , as well as prototyping and management interface on the Android platform.

%[David]
The conceptual framework proposed at this work was tested through a reference implementation called \mbox{SmartAndroid}~\footnote{For more information visit www.tempo.uff.br/smartandroid}. The project includes a programming API to help the developers in constructing UbiApps as well as offers basic management services to provide resource registration, discovery and localization. The SmartAndroid has been developed over Google Android platform as we believe that in a near future it will be widely used in sensors and devices embedded in furnitures and home appliances.
% [we have implemented most of the framework features, thought much more remain to be done]

%proposta de uso de marketplace
% A exposição de dados por meio de \textit{Content Provider} permite a integração de aplicativos externos ao servidor. Estes aplicativos podem ser dispostos em um \textit{marketplace} (http://developer.android.com/distribute/open.html) permitindo que tanto usuários comuns quanto programadores incrementem o ambiente com novas aplicações ubíquas. 
The SmartAndroid middleware contemplates the possibility of running multiple applications in the same IE. Thus, focusing on the user and inspired by the Google Play -- which is a application market for Android devices --, we have as a future work the creation of a SmartAndroid Market~\cite{google2013}, where users will be able to easily download and start different applications. Also similar to the Google Play, a Manifest file will require permission for accessing the domains, what will allow users to maintain control over what off-the-shelf applications have permission to know and to do. Within the market, the UbiApp could be tested towards discovering any possible harm that the application could do and any possible inefficiency that could consume too much the energy resources of the host device.
% TODO: talk about market place
% [security concern on store: helps the discovery of malware before the user download]

  \subsection{Location}

  [talk about location]

  The devices' indoor localization may be acquired through many different approaches, as an example, using radio frequency devices such as RFID tags and Wifi devices~\cite{Talukder2007Tracker} or even a smart floor~\cite{Branzel2013SmartFloor} which can locate people and objects by a pressure-sensing floor. However, the SmartAndroid doesn't implement a specific technique to locate resources, in order to keep the framework generic. The SmartAndroid API deal with location in a hi-level way, making the data acquirement transparent to it. This aspect is important to keep the system flexible to any approach, allowing the developer to use different implementations.

  \subsection{Communication Support}
  %[talk about stub intantiation calls]
  %TODO: example of CV and OP? Maybe it is worth add some on the text
  To perform any interaction with a RA that is not locally instantiated, the UbiApp must use a stub (also called proxy) that is bound to the referred RA. This stub implements the same interface as the RA, i.e. the same CVs and OPs, which are methods, from a Object-Oriented Programming perspective. The stubs could also be automatically generated since their role is to forward the calls to the implemented RA. With the stub on hands, the application can invoke remote methods, as if they were locally invoked. These remote methods may be to retrieve a CV or to trigger an OP. The Figure~\ref{fig:rpc} represents this direct, or synchronous, communication. The steps to achieve a direct communication, as represented by the figure, are as follows:
  \begin{enumerate}
    \item The UbiApp calls a method from the stub 
    \item The call is forwarded to the RA instance through the network in the second step
    \item This instance can interact with the device (the lamp, in case)
    \item The result or the acknowledge is then returned to the caller
  \end{enumerate}

@@

  \subsection{Context Rules Composition} \label{subsec:rules}
%   [implementation of context rules]
  The SmartAndroid supports the context interpretation by providing mechanisms to create context rules. The context rules life cycle has two major phases: composition and running. The composition phase corresponds to the instantiation of a CI and of an actuator and the binding of both through subscription. The running phase is when the CI and the actuator(s) are already instantiated as the other RAs that feed the IE with context information, and they are able to exchange messages and perform their job.
  %   Support the interpretation of the context framework enables a separation of concerns, where developers abstract implementation of contextual rules and focus on the application logic.

  The composition of context rules can be done basically in two ways. The former is at the implementation level, where developers are able define their context rules in a description file or using a programming language. The latter is at end user level, where non trained people are able to define their context rules supported by the PIPA (Subsection~\ref{subsec:ipgap}).

  At the implementation level, a context rule can be described in a JSON structured file (or byte stream) following a specific format. In the first part an expression must be defined and in the second part the actuators can be defined. This expression contains conditions that are the comparison of RA's CVs with constant values (e.g., Stove.OvenTemperature = $180\,^{\circ}\mathrm{C}$), of two different CVs from the same or from different RAs (e.g., ArConditioning.Temperature $>$ OutsideSensor.Temperature), or the RA's CV alone if it is has a boolean value (e.g., Stove.IsOn). The conditions are bound to each other through the logical operators ``AND'' and ``OR'', it is also allowed the use of ``NOT'' and of parenthesis, what makes possible to compose more complex expressions by using subexpressions. The expression or subexpression may also contain a timer.

  The definition of actuators is simpler than the definition of the CI and is not mandatory, since other actuators can be bound to the IC later. A actuator is composed by a list of actions, where each action have a RANS, a OP reference and the parameters (e.g., Stove.SetOvenTemperature to $100\,^{\circ}\mathrm{C}$). Both the actuator and the CI have a name, that makes easier to search for them.

  The SmartAndroid has a parser that can read the JSON file and create the instances of the CI and actuators. After reading the rule expression the parser creates a tree data structure that corresponds to that expression in memory, and the timers are also referenced within the structure. The CI created queries all the CVs that appear on the expression to keep a cache of the context information and to start the evaluation, next the CI subscribes to all these CVs and sets itself as ``running''. After the creation of the CI, the parser creates instances for the actuators, which are subscribed to the CI previously instantiated.

  % [context rules conflicts is an ongoing study that aims at discover a conflict in runtime,]
  A common concern when having multiple context rules running in parallel is the occurrence of conflicts. The conflict happens when two CIs may be validated at the same time and the actuators associated with them have opposite actions. The conflict has also happen when a context rule causes the triggering of another that may cause the triggering of the first one, leading to a circular deadlock. We have as a future work to focus on this concern and to provide a solution for the detection and resolution of conflicts.
%   , for instance:
%   % \vspace{-2mm} 
%   \begin{itemize}
%   % \begin{compactitem}[$\circ$]
%   \item IF stove is on and oven temperature $>$ 100 for 15min, then lower the temperatura;
%   \item IF stove is on and oven temperature $<$ 180 for 15min, then raises the temperature.
%   % \begin{compactitem}
%   \end{itemize}
%   % \vspace{-1.5mm}

%   The detection and resolution of conflicts aims at identifying the creation of contradictory rules and at providing a solution. The solution to this type of conflict involves the comparison of the expressions of a new IC with the existing (at the time of connection to the actuator) and, if found expressions that intercede, check the actions of the actuators to see if they are the same type but with different inputs.

%   A multiplicação de regras de contexto no AmbI, inclusive por diferentes usuários e aplicações, pode facilmente levar a casos de conflito. O conflito ocorre quando dois ICs possuem expressões que se intercedem enquanto os respectivos atuadores possuem ações opostas. Por exemplo, considere as seguintes regras:
%   % Neste caso, o fogão não foi ligado e esquecido, mas sim ligado para assar um bolo. 
%   O usuário criou uma regra para controlar a temperatura do fogão, quando já havia outra cadastrada que tinha ação oposta mas que poderia ocorrer ao mesmo tempo. As regras estão, portanto, em conflito, pois não está claro como o sistema deve agir.
%   %FIXME: compactar itemize
% 
%   % [ SOLUÇÃO PARA RESOLUÇÃO DE CONFLITOS ]
%   A detecção e a resolução de conflitos visam identificar a criação de regras contraditórias e prover uma solução, atendendo ao requisito de confiabilidade. A solução para este tipo de conflito envolve a comparação das expressões de um novo IC com as já existentes (no momento da ligação com o atuador); e, caso encontradas expressões que se intercedam, conferir as ações dos atuadores para ver se são do mesmo tipo mas com diferentes entradas.

@@

  %[David]
  The main purpose of the PIPA is provide to developers a hybrid environment where simulated and real resources can interact to each other transparently, enabling their users to construct UbiApps as well as test and evaluate them through a wide variety of scenarios. Thus RAs from sensors, actuators and household appliances can be graphically represented at our tool, along with a schematic map from the intelligent environment so that some of the resources aspects as its context and location can be easily visualized on the map, helping in application debugging.

  Besides, PIPA operate with two different interfaces: the \textit{developer interface}, already mentioned, and the \textit{end-user interface}. In the end-user interface, users in general, i.e. non-technical users, are able to use PIPA to manage their intelligent environment remotely using a mobile device, such as a tablet. Through an intuitive GUI, end-users can not only visualize the state of their home devices but control them. For instance, it's possible turning off a stove burner, change the air-conditioner temperature, and so on. By touching the device icon in the screen, an interface that presents the device's main functions is shown, where the user can operate it as the tool were a remote control.

  % [main figure]
  \begin{figure}[ht]
  \centering
  \includegraphics[width=3.5in]{figures/map-forniture-cv-menu-big-v2.pdf}
  \caption{Management and Prototyping Interface} %PIPA
  \label{fig:ipgap}
  \end{figure}

%   [rules composer]
  %TODO Think about a better rule example!!
  In addition, end-users can compose context rules in a hi-level way by graphically combining the CVs from any RA and logical connectives, forming logical expressions as \textit{``IF the TV from living room is on, AND nobody has been present at the living room for 15 minutes, THEN turn off that TV''}. To compose this rule the user have to, first of all, select the rule composer mode in the PIPA. Then, he must touch the TV icon at the living room representation on the map, and choose the desired CV -- in that case, the CV that refers if the TV is turned on or not. After that, a dialog box showing all possible logical operators is shown, so that the user can select one of them as the ``AND'' operator. The user can repeat this process, adding others CVs and operators until the expression is complete. Finally, the rule can be associated to one or more actions, provided by actuators. When the context rule is finished, the SmartAndroid writes down this rule in a JSON file respecting the standard, thus if the system crashes or terminates, all the composed rules can be afterwards re-started.

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%              SECTION
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Example Applications} \label{sec:evaluation}
Some applications have been built to explore the potential of the framework for building UbiApps. Although the example applications are simple, and alone does not justify the deploy of the SmartAndroid in the environment, they have allowed us to validate the features provided by the framework and achieved by the SmartAndroid. It also demonstrated the possibility of concomitant running of UbiApps, what similar to an operation system where different processes run separately but in the same system. Within the SmartAndroid, UbiApps run separately but in the IE.% and can share resources,

% In this section we will elaborate over the SmartLiC application, and next we will briefly talk about the others.
In the subsections that follows we will elaborate over some UbiApps that have been developed, or are still in development, in the course of the project. Besides these applications we have also implemented some context rules, such as the `Forgotten Stove'' that have appeared throughout the text.
% the SmartLiC, the MediaFollowMe, remote control type applications where resources are controlled remotely by another resource (in this case, a smartphone ), among other applications. A future work is to install a system in Ambi home care health, as in ~ \ cite {} carvalho2010sciads.

@@

  The PIPA has allowed us to test this application. We instantiated lamp simulators for each room and an avatar for a user, and also a simulator for day light sensor. Next context rules were created for each room of the house. The following enumeration shows the initialization routine of the SmartLiC UbiApp.
  %Next context rules were created by filling a skeleton of a rule written in json, what caused the instantiation of a CI and actuator for each room.  If the avatar is dragged, in the PIPA, from a room and dropped in another, and the specified time passes, the the lamp is turned off.
  \begin{enumerate}
   \item Create a context rule skeleton in JSON
   \item Get the rooms from the house
   \item Get the lamp and presence sensor of each room and fill the rule expression skeleton
   \item Instantiate a CI with the rule expression
   \item Instantiate a actuator to perform the actions declared in the context rule skeleton
   \item Bind the CI and the actuator
   \item Repeat this for each room
  \end{enumerate}

@@

  The SmartLiC allowed us to validate two features of the RLS, which are the discovery of the rooms and the RAs of some specific type are within the room (i.e. the lamps and presence sensors). In this UbiApp we also have created a set of context rules with timers using the context rules parser (including CI and actuators). And the PIPA were well explored in this implementation, since we has shown the triggering of many events in parallel and multiple users cohabiting the same IE.
% 
%   After the start of the context rules, 
%   \begin{enumerate}
%    \item On event received, get the sender, the 
%   \end{enumerate}

@@

  \begin{figure}[ht]
  \centering
  \includegraphics[width=3.5in]{figures/beagle-lamp.pdf}
  \caption{Remote Control} %PIPA
  \label{fig:remotecontrol}
  \end{figure}

@@

  This game was implemented with simulated resources and using the PIPA. The players were represented with avatars and their location were updated with the drag and drop of the avatars over the map.

@@

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%              SECTION
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion} \label{sec:conclusion}
% [framework main features]
In the present paper we have proposed a framework that address some major issues that are found in the construction of ubiquitous applications, both for developers as for end users. We have focused on three challenges, that have guided the development of the framework. For dealing the the heterogeneity of devices present in the intelligent environment we have created the Resource Agent abstraction that encapsulates the mechanisms to interact with physical and also virtual resources, and includes a synchronous and asynchronous communication mechanism. The variety of context information and services is achieved by the definition of Context Interpreters that are responsible for aggregating context information, providing it in a higher level. And for the resource availability issue we have created the PIPA, which is a software that helps the prototyping and testing of ubiquitous applications through a easy to use interface, and allows the visualizations of applications running in a environment populated with real and simulated devices.

@@

% [future work]
% [more complex resource discovery]
% [aggregate other security concerns]
% [machine learning and data mining]
% [internet of things - expand to internet]
As future work we plan to extend the framework to allow more complex queries in the resource discovery service, what nowadays is still strict; adapt the communication mechanism for using web services, and get our approach closer to the concept of the internet of things; further research on other security concerns; provide mechanisms to avoid/solve potential conflicts on context rules; aggregate machine learning techniques that will enable a better understanding of user preferences, and will make possible the offering of context rules and other helping services; improve the PIPA and evaluate its usability, using methodologies of human-machine interface; and son on.
